### 2020-09-09 paper.tex
Major edits in paper, reimplemented analysis in Jupyter

% GigaScience template
\documentclass[a4paper,num-refs]{oup-contemporary}

\journal{gigascience}
%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{csvsimple}
\usepackage{minted} % Used for JSON highlighting
\usepackage{datatool}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{caption}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{anyfontsize}
\usepackage[flushleft]{threeparttable}
\usepackage{xspace}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\change}[2]{\color{cyan}Changes: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\rom}[1]{\lowercase\expandafter{\romannumeral #1\relax}}
\newcommand{\tristan}[1]{\color{blue}From Tristan: #1\color{black}}

\renewcommand{\labelitemi}{$\textendash$}
\title{Security Vulnerabilities of Container Images in Neuroimaging}
\begin{filecontents*}{resul.csv}
labels,names,A,C,V,Stools
a,bids/example:latest,838,663,683,
b,bids/freesurfer:latest,353,215,192,
c,bids/ndmg:latest,934,757,782,
d,bids/broccoli:v1.0.0,203,203,117,
e,bids/fibredensityandcrosssection:latest,1086,849,868,
f,bids/spm:latest,191,246,235,
g,poldracklab/mriqc:0.15.0,432,328,332,
h,bids/cpac:v1.0.1a\_22,698,524,535,
I,bids/hyperalignment:latest,1717,1327,1433,
j,bids/mindboggle:0.0.4-1,671,955,569,
k,bids/mrtrix3\_connectome:latest,275,157,158,
l,bids/rs\_signal\_extract:0.1,867,454,473,
m,bids/aa:v0.2.0,189,244,233,
n,bids/niak:latest,1482,1366,1420,
o,bids/oppni:v0.7.0-1,913,698,723,
p,poldracklab/fmriprep:unstable,516,355,393,
q,bids/brainiak-srm:latest,897,484,503,
r,bids/nipypelines:0.3.0,641,878,451,
s,bids/hcppipelines:v3.17.0-18,526,873,620,
t,bids/magetbrain:latest,112,117,126,
u,bids/tracula:v6.0.0-4,874,664,683,
v,bids/baracus:latest,172,238,216,
w,bids/antscorticalthickness:v2.2.0-1,289,27,no data,
x,bids/dparsf:v4.3.12,507,750,757,
y,bids/afni\_proc:latest,67,16,no data,
z,bids/rshrf:1.0.1,5,3,3,
aa,bids/ndmg:v0.1.0,934,758,782,
ab,mcin/qeeg:latest,121,121,142,
ac,camarasu/creaphase:0.3,148,148,186,
ad,bigdatalabteam/hcp-prefreesurfer:exec-centos7.freesurferbuild-centos4-latest:latest,257,266,157,
ae,mcin/docker-fsl:latest,178,178,201,
af,gkiar/onevoxel:v0.3.0rc2,5,2,2,
ag,bigdatalabteam/hcp-prefreesurfer:exec-centos7-fslbuild-centos5-latest:latest,257,266,157,
ah,bt5e/ants:latest,592,600,336,
ai,poldracklab/fmriprep:1.2.3,591,441,447,
aj,gkiar/dwipreproc\_fsl-5.0.11\_minified:latest,391,258,263,
ak,gkiar/mask2boundary:v0.1.0,15,11,11,
al,mcin/ica-aroma:latest,178,178,202,
am,c3genomics/genpipes,,,,99
an,bioinformatics-group/aqua-singularity-recipe,,,,7
ao,MontrealSergiy/BEst,,,,106
ap,shub://aces/cbrain-containers-recipes:fsl\_v6.0.1,,,,206
aq,shots47s/bids-freesurfer-6.0,,,,212
ar,shots47s/bids-fmriprep-1.2.3,,,,453
\end{filecontents*}

\begin{document}

\author[1]{Bhupinder Kaur}
\author[1]{Mathieu Dugr\'e}
\author[1]{Aiman Hanna}
\author[1]{Tristan Glatard}

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}

\maketitle

\begin{abstract}
Software containers have become key technologies for the deployment of
scientific data analyses on high-performance computing (HPC)
infrastructures, due to their positive impact on the portability, deployment, and
reproducibility of computations. Container images, however, often contain slightly outdated
software packages compared to HPC environments maintained by professional
 system administrators, and they often carry more
software packages than required by analyses, including compilers or unused
system librairies. As a result, container images are likely to contain
important amounts of security vulnerability, increasing the attack surface
for the entire infrastructure. This paper presents a vulnerability analysis
of container images for scientific data analysis. We analyze container
images with four vulnerability scanners, focusing on the use case of
neuroscience data analyses. We quantify the effect of images updates and
size reduction on the number of vulnerabilities, and conclude with
recommendations on how to build container images with a reduced amount of
vulnerabilities.

\end{abstract}

\begin{keywords}
Containers; Docker; Singularity; Security Vulnerabilities; Neuroimaging.
\end{keywords}

\section{Introduction}

% Containers are popular on HPC, in particular Singularity
Software containers have emerged has an efficient solution to deploy
scientific data analyses on High Performance Computing (HPC) clusters, due
to their portability, ease of use, and limited overhead. The
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
Docker\footnote{\url{http://docker.com}} for HPC deployments, due to its
secure handling of multi-user environments, and its convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the globe, and used routinely for Big Data analyses.

% Containers are meant to isolate applications from the host, but security
% remains an issue
Leveraging core features of the Linux kernel such as namespaces, control
groups and chroot, containers isolate processes from the host computer,  
and control the memory, CPU, network and file-system resources assigned to
them. Nevertheless, containers still share the kernel, mounted file systems
and some devices with the host, which raises security
issues~\cite{martin2018docker, sultan2019container, combe2016docker} and
opens the door to privilege escalation, denial of service, information
leak and other types of attacks~\cite{gantikow2016providing}.

% Container images, our study
Container images often include nearly complete operating-system
distributions in addition to the data analysis software and its required
dependencies. Images are rarely updated, as they are often built once for
the entire duration of an analysis, to avoid results perturbations due to
software updates~\cite{gronenschild2012effects,
glatard2015reproducibility}. Images also commonly include more dependencies
than required by a specific analysis, to facilitate reusability across
experiments. As a result, it has been shown that over 30\% of official
images in DockerHub contain high priority security
vulnerabilities~\cite{gummaraju2015over}, that both official and community
images contain more than 180 vulnerabilities on average~\cite{Shu2017}, and
that vulnerabilities are often due to outdated
packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerability analysis of container images
used in HPC clusters.  
Our research questions are the following:

\textit{RQ1: What is the current amount of known vulnerabilities in
container images deployed on HPC clusters?} Vulnerabilities are possible
attack vectors that may seriously compromise the security of HPC clusters
and the integrity of user data. We report vulnerability analyses produced
by four popular image scanning tools, Anchore, Vuls, Clair, and Stools.

\textit{RQ2: Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and the life-cycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the vulnerability
analysis of RQ1.

\textit{RQ3: Can the amount of vulnerabilities be reduced by minifying the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

We answer these questions by taking the particular example of neuroimaging,
a domain that has entered the Big Data era as many other
ones~\cite{van2014human}. The next section describes the materials and
tools used in the experiments, including container images and image
scanners. We then describe our image update and minification methodology to
reduce vulnerabilities. Results present the detected vulnerabilities in
container images, explain observed discrepancies between scanners, and
report on the efficiency of image update and size reduction. As a
conclusion, we provide a set of image creation guidelines for a more secure
deployment on HPC clusters.

% We focus on the analysis of container images
% present in BIDS apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}, two containerized
% application frameworks commonly used in neuroimaging.

\section{Materials and Tools}

To detect security vulnerabilities
in container images and examine the effect of different
manipulations on these images, we used container
images from two popular application frameworks, as well as
four of the main image scanners.

\subsection{Container Images}

We scanned all container images available at the time of this study on two containerization frameworks
used in neuroscience: BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques},
totalling
44 container images.
Out of these 44 images, 26 were obtained from BIDS apps 
and 18 from Boutiques. At the time of the study, BIDS apps had 27 images,
out of which one wasn't available on DockerHub. Boutiques had 49 images,
%listed using Boutiques \texttt{bosh search} command,
however, only 23 unique images were listed, out of which 3 couldn't be retrieved and 2
were already included in BIDS apps. All the final 26 images
from BIDS apps were Docker images, whereas Boutiques contained 18 Docker images
and 6 Singularity images.

\subsection{Image Scanners}

We used four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair~\cite{arminc_2019} and
Singularity Container Tools --- Stools~\cite{stools}. Anchore, Vuls and Clair support Docker
images while Stools support Singularity.

Anchore is an end-to-end, open-source container security platform. It
analyzes container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
In our experiments, we use Anchore Engine version 0.5.0 through Docker image \texttt{anchore/anchore-engine:v0.5.0}, and
Anchore vulnerability database version 0.0.11.

\begin{figure*}
\begin{tabular}{|c|c|c|c|}
 \hline
\textbf{OS} &	\textbf{Anchore} &	\textbf{Vuls} &	\textbf{Clair} \\
\hline
	\textbf{Alpine} & Alpine-SecDB &	Alpine-SecDB &	Alpine-SecDB \\
\hline
	\textbf{CentOS} & Red Hat OVAL Database & Red Hat OVAL Database and Red Hat Security Advisories & Red Hat Security Data \\
\hline
	\textbf{Debian} & Debian Security Bug Tracker &	Debian OVAL Database and Debian Security Bug Tracker & Debian Security Bug Tracker \\
\hline
	\textbf{Ubuntu} & Ubuntu CVE Tracker &	Ubuntu OVAL Database &	Ubuntu CVE Tracker \\
 \hline
\end{tabular}
\caption{Vulnerability databases used by scanners for different OS distributions}
\label{database}
\end{figure*}

Vuls is an open-source vulnerability scanner for Linux and FreeBSD. It
offers both static and dynamic scanning, and both local and remote
scanning. In our experiments,we used Vuls 0.9.0, executed through Docker image
\texttt{vuls/vuls:0.9.0} in remote dynamic mode.

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
scans Docker images layer by layer and maintains a database of vulnerabilities. We used Clair through Clair-scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. 
% Clair-scanner scans the image,
% prepares a list of vulnerabilities, compares that list against a
% whitelist, and flags vulnerabilities that are not present in the whitelist.
% In our experiments, we did not use a whitelist to filter scanning results in
% order to make a fair comparison between scanners.
% Clair-scanner maintains a Docker image with the up-to-date vulnerability
% database from a set of different sources.
% Figure~\ref{database} lists
% vulnerability databases that Clair referred for Alpine, Ubuntu, Debian, and Centos.
% Clair also uses NVD database for getting vulnerability metadata.
We used Clair version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

Singularity Tools (Stools)
are an extension of Clair for Singularity images. Stools
exports Singularity images to \texttt{tar.gz} format, acting as a single layer Docker image
to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we used Singularity Tools version 3.2.1 through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}.
% , which is using another Docker
% image \texttt{arminc/clair-db:latest} for referring to vulnerability feed.
Since Stools uses Clair internally for scanning, vulnerability databases used
by Stools are same as mentioned for Clair.
To scan Singularity images, we followed the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github}.

\subsection{Vulnerability Databases}

Scanners refer to two types of vulnerability databases. The first one is
the Open Vulnerability and Assessment Language (OVAL) database, an
international open standard that supports various OS distributions
including Ubuntu, Debian and Centos but not Alpine. The second one are vulnerability
databases from specific OS distributions, such as Alpine-SecDB, Debian
Security Bug Tracker, Ubuntu CVE Tracker, or Red Hat Security Data. OS
distributions often assign a status to each vulnerability, to keep track of
required and available security fixes in different versions of the
distribution. Figure~\ref{database} shows the vulnerabilities databases used by each scanner 
by OS distribution. Vuls uses OVAL databases for all distributions except Alpine, since Alpine is not available in OVAL. On the contrary, 
Clair exclusively refers to distribution-specific databases. Anchore uses OVAL only for CentOS \tristan{why?}. It is also worth noting that there is
no vulnerability data for Ubuntu 17.04 and 17.10 distributions in the OVAL database, since these distributions have reached end of life, meaning that images
with these distributions cannot be scanned with Vuls \tristan{why doesn't vuls use Ubuntu CVE tracker?}.
\tristan{In Vuls, what happens when 2 databases are used?}

Anchore also refers to National Vulnerability Database (NVD) to get
metadata of vulnerabilities. Apart from this, like Anchore, Vuls also
refers NVD database for getting vulnerability metadata \tristan{same
question}. \tristan{why is it not in the table?}.

For Centos images, Anchore and Clair gives scanning results using Red Hat
Security Advisory (RHSA) identifiers for Centos images, whereas Vuls uses
CVE identifiers. RHSA identifiers specifies security patches that are
pending to apply and CVE identifiers specifies vulnerabilities that exist
in the system. We mapped RHSA identifiers to corresponding CVE identifiers,
to allow for a comparison between scanners.

Different vulnerabilities may be
reported by scanners if scanning experiments take place on different dates.
To avoid such discrepancies, we froze the vulnerability
databases used by these scanners as of 2019-09-25, when we initiated our
experiments.\tristan{this contradicts the date used for Clair}

\subsection{Image Update}

A first approach to reduce the number of vulnerabilities in container
images is to update their packages to the latest version packaged by the OS
distribution. To study the effect of such updates, we developed a script
(available
\href{https://github.com/big-data-lab-team/container-vulnerabilities-paper/blob/master/Scripts/update}{here})
to identify the package manager in the image, and invoke it to update all
OS packages. \tristan{what was the date at which update was run?}

\subsection{Image Minification}

%We revert to image minimization to reduce the number of vulnerabilities without affecting results reproducibility.
%An interesting choice to do so would be to use 
%Neurodocker~\cite{neurodocker}.
%This tool traces and deletes unnecessary files.
%However, it does not affect the number of installed packages.
%Since image scanners report vulnerabilities according to the installed
%packages. Therefore, after
%minimization the same number of vulnerabilities is still reported.
%Consequently, we need to reduce the number of installed packages
%in the images by deleting unnecessary packages. For that, it is required to figure
%out which packages we need for the pipeline to function properly.

A second approach to reduce the number of vulnerabilities in the images is
to remove unnecessary packages, an operation potentially specific to each
analysis. We used the open-source \reprozip tool~\cite{rampin2016reprozip}
to capture the list of packages used by an analysis. For this purpose,
\reprozip first captures the list of files involved in the analysis,
through system call interception, then retrieves the list of associated
software packages, by querying the package manager. We extend this list
with a passlist of packages required for the system to function
\tristan{can you give an example?}, and with all the dependencies of the
required packages, retrieved using Debtree~\cite{debtree} (Repoquery
(\url{https://linux.die.net/man/1/repoquery}) could be used in RPM-based
distributions instead). Our minification script, available
\href{https://github.com/big-data-lab-team/container-vulnerabilities-paper/tree/master/Scripts/minification}{here},
installs \reprozip in the image to minify, runs an analysis to collect a
\reprozip trace, and finally deletes all unnecessary packages.

Using this approach, we minified five Debian- or Ubuntu-based BIDS app images,
using basic analysis examples found in the applications documentation.

% \subsection{Hybrid Approach}

% Following the two previous approaches, we employ a hybrid approach, where
% we combine both update and size reduction processes in an attempt to
% lessen number of vulnerabilities or to test the effect of
% combining two approaches. To allow for that, we first update all packages in
% the image, then delete all unnecessary packages. Applying these processes in reverse
% may reduce the needed work; however this is not possible in all cases. For instance,
% some packages, such as \texttt{apt-transport-https}, are required to fetch updates, deleting these packages
% ahead of time will disallow the update process to take place.

\section{Results}

Figure~\ref{fig:vulnerabilities} summarizes our results. All the collected
data are available in our GitHub repository at
\url{https://github.com/big-data-lab-team/container-vulnerabilities-paper}
with a Jupyter notebook to generate the figures.

\subsection{Detected Vulnerabilities}

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/figure.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected
by Anchore in container images. \textbf{(A)} Number of vulnerabilities
by container image and severity. \textbf{(B)} Combined
effect of image minification and package update on 5 container images. \textbf{(C, D)}
Number of vulnerabilities by number of packages, before and after
package update.}
\end{figure*}

A substantial number of vulnerabilities were found in the tested container
 images (Fig~\ref{fig:vulnerabilities}-\textbf{A}), with an average of 466 vulnerabilities
  per image and a median of 372. In comparison, no vulnerabilities were found in base
 Docker images \texttt{ubuntu:20.04} and \texttt{centos:7} after package
 update. Moreover, a significant fraction of detected vulnerabilities are
 of high severity
 (\href{https://www.first.org/cvss/specification-document}{CVSS} score
 >=7.0) and a few of them are of critical severity (CVSS >= 9.0). Remote
 attackers could possibly exploit these vulnerabilities to execute
 arbitrary code in the container, by crafting responses to specific network
 requests. The OS distribution is reported only for illustration purposes,
 as no obvious association between the OS distribution and numbers of
 vulnerabilities could be found.

Unsuprisingly, a strong linear relationship is found between the number of
detected vulnerabilities and the number of packages present in the
container image (Fig~\ref{fig:vulnerabilities}-\textbf{C}, r=0.83,
$p\textless10^{-10}$). On average, 1.8 vulnerabilities are introduced for
each new package installation. This observation motivates a systematic
review of software dependencies by application developers, to avoid
unnecessary packages in container images. This is also an argument in favor of lightweight
distributions such as Alpine. Compared to Ubuntu and Debian distributions,
CentOS images seem to have a lower number of vulnerabilities by package on
average, although data is too scarce to conclude.

\subsection{Effect of image udpate}

Updating container images reduces the number of vulnerabilities by package
by a factor of 3 on average, leading to an extra 0.6 vulnerabilities by
package (Fig~\ref{fig:vulnerabilities}-\textbf{D}, r=0.81,
$p\textless10^{-7}$). Six container images are missing on this figure; they
could not be updated since their OS distributions reached end of life
\tristan{give more details, one of these images is centos 7 which didnt
reach EOL}. In spite of the associated reproducibility challenges, updating
packages therefore appears as a very good way to avoid vulnerabilities. It
is not an ultimate solution though, as substantial numbers of
vulnerabilities remain.

\subsection{Effect of minification}

Another approach to reduce the number of vulnerabilities involves deleting
unnecessary packages from the container images. It is a tedious operation,
as it requires running an actual data analysis in the container image, to
identify the packages required by the application. In addition, the
resulting container image is only valid for the specific type of analysis
used in the minification process, as other executions might require a
different set of packages.

Using the ReproZip-based approach described previously, we minified 5
different images covering the spectrum of detected vulnerabilities
(Fig~\ref{fig:vulnerabilities}-\textbf{B}). We find that minification reduces the
number of vulnerabilities, albeit less systematically than package update.
For some container images, such as image \textbf{S}, minification removes more
than 70\% of the detected vulnerabilities. For other images, such as
image \textbf{f}, it only reduces the number of vulnerabilities by less than 1\%.
The effect of minification stems from the number of packages
than can be removed, which varies greatly across images. For
instance, images \textbf{f} and \textbf{a} have a large number of packages,
but the last majority of them is required by the analysis, which makes
minification less useful. In other cases, a limited number of unnecessary packages contain 
a significant number of vulnerabilities, which makes minification very impactful. 
This was the case in images \textbf{d}, \textbf{S} and \textbf{U}, where removing compilers
and kernel headers reduced the number of vulnerabilities by an important fraction.

\subsection{Combined effect of image update and  minification}

Package update and image minification remove different types of
vulnerabilities. The former is efficient against vulnerabilities that have
been fixed by the package maintainers, while the latter targets unused
software. In four of the five tested images, we find that combining update
and minification further reduces the number of vulnerabilities compared to
using only one of these processes
(Fig~\ref{fig:vulnerabilities}-\textbf{B}) \tristan{revise when
minification in image f is clarified}. The combined effect of update and
minification seems to be dependent mostly on the effect of minification: if
minification is useful, then it is beneficial to run it in combination with
package update.

\subsection{Discrepancies between scanners}

The results presented so far were obtained with Anchore, a popular
container image scanning tool. We scanned the images with two other tools,
Clair and Vuls, to evaluate the stability of our results.

% \begin{table*}
% 	\csvreader[%
% 	 tabular={|c|c|c|c|c|c|},
% 		table head = \hline\textbf{Labels} &\textbf{{Image names}} & \textbf{Anchore} & \textbf{Clair} & \textbf{Vuls} & \textbf{Stools}\\\hline,
% 		late after line= \\,
% 			late after last line=\\\hline %
% 		]{resul.csv}{labels=\labels,names=\names,A=\A,C=\C,V=\V,Stools=\Stools}%
% 			{\labels & \names & \A & \C & \V & \Stools}
% 		   \centering
% 		\vspace*{1mm}
% 		\caption{\label{table1}Number of vulnerabilities detected by the four scanners. Vuls shows \textit{no data} for images \texttt{bids/antscorticalthickness:v2.2.0-1} and \texttt{bids/afni\_proc:latest}
% 		since their OS distributions reached end of life.} 
% 	\end{table*}
	\begin{figure}
        {\includegraphics[scale=2.5,width=\columnwidth]
        {Figures/venn.pdf}}
	\vspace*{-10mm}
	\caption{\label{fig:venn} Scanner results}
\end{figure}

% Table~\ref{table1} shows the number of vulnerabilities detected in the different images
% by the four scanners. 
% \change{Our experiments concluded considerable discrepancies in the results obtained by
% these scanners. %We higlighted all Centos images in the Table~\ref{table1} since.
% While comparing scanner results, we exluded two images \texttt{bids/antscorticalthickness:v2.2.0-1} and
% \texttt{bids/afni\_proc:latest} because Vuls
% cannot scan these image, hence no results are present to compare.
% %The reason behind excluding
% %vulnerability identifiers for Centos. Anchore and Clair use Red Hat Security Advisories (RHSAs)
% %to report pending patches that need to be applied to specific packages,
% %whereas Vuls uses Common Vulnerabilities and Exposures (CVE) number to report
% %vulnerabilities in packages. We exluded other two images because Vuls
% %can not scan these image, hence no results are present to compare.
% In most cases,
% Anchore reports more vulnerabilities than Vuls and Clair resulting in 
% the Jaccard coefficient of 0.63 with Clair, 0.59 with Vuls,
% whereas Vuls and Clair share more common vulnerability detection resulting to the Jaccard coefficient of 0.86.
% Figure~\ref{fig:venn} represents the number of vulnerabilities reported by
% the scanners in Docker images and the overlap between them.
% This Figure also excludes two non-comparable images, which cannot be scanned with Vuls.
% It is interesting to note that out of all reported vulnerabilities, 4453 vulnerabilities are only
% reported by Anchore, 699 by Clair, and 536 by Vuls.}
% \newline\\
% As the number and type of vulnerabilities reported by the these scanners are significantly different from each other,
% we desire to investigate the reasons behind these differences.
% Following our investigation, we are able to explain some of the reasons behind
% the differences. 
% Such reasons were resulted from the way these scanners function or behave 
% in particular, the following actions by the scanners are indicative:
% \begin{itemize}
%    \item scanners ignoring vulnerabilities in Linux-kernel-header packages,
%    \item scanners referring to different vulnerability databases,
%    \item different frequency of updating their databases,
%    \item bugs in scanners. 
% \end{itemize}

% Below, we explain these reasons in further details and provide examples
% for clear illustrations.

% \textbf{Linux-kernel-headers:} Clair and Vuls ignore all vulnerabilities in the Linux-kernel-headers,
% whereas Anchore reports all of them.
% We find that 4443 out of 4453 vulnerabilities that are only reported by Anchore are present in Linux-kernel-headers.
% For example, CVE-2019-9506 present in Linux-libc-dev-4.4.0-31.50, reported by Anchore, is not
% detected by both Vuls and Clair.
% Linux-kernel-headers are the header files that only provide function names, their parameters, and their
% return types.
% The actual source code of those functions is present in the Linux kernel and is not inside the image.
% So, vulnerability detection in header files
% is irrelevant in this case.

% \textbf{Ubuntu vulnerabilities with status as “Not-affected”:} Vuls reports vulnerabilities that have status
% as "not-affected" in the Ubuntu database. It means that according to Ubuntu the image is not affected
% by the particular vulnerability. In other words, Vuls reports vulnerabilities that it should not have been reported.
% For example, when we scan an image having the Trusty release in it and Vuls reports CVE-2015-2305
% in the package Cups. However, upon checking the Ubuntu vulnerability database, we find that package Cups in Trusty is not
% affected by this vulnerability.

% \textbf{Ubuntu vulnerabilities with status as "Ignored (reached end-of-life)":}
% Some vulnerabilities are ignored in Ubuntu 14.04 LTS as it reached end-of-life. 
% However, in Ubuntu 14.04 ESM release, these vulnerabilities may or may not be present.
% Anchore checks the last status of these vulnerabilities in LTS whether the package was vulnerable
% prior to the end-of-life. Only if it was vulnerable, Anchore reports it, otherwise it does not.
% If this kind of vulnerability is present in ESM release then Clair reports it because it
% only refers to ESM release. On the other hand, if the vulnerability
% status in ESM release is "DNE" (Does Not Exist), then Clair does not report it. 
% For example, CVE-2017-9994 vulnerability in Libav package is ignored in Trusty LTS release due to its end-of-life.
% Consequently, it is not reported by Anchore. This vulnerability has status as "DNE" in Trusty ESM
% release so Clair also does not report it. As Ubuntu 14.04 LTS release is present inside the images,
% Clair's referring to Ubuntu 14.04 ESM release vulnerability database is incorrect.

% \textbf{Vulnerabilities not present in Ubuntu 14.04 ESM release:} These vulnerabilities are not present in Ubuntu 14.04 ESM so they 
% are not detected by Clair but they affect Ubuntu 14.04 LTS, and hence detected by Vuls and Anchore. The CVE-2017-13758 is an example
% of such vulnerabilities.

% \textbf{Misses:} Clair and Vuls
% missed vulnerabilities that they should have detected. For example, the CVE-2016-9082 
% is present in package Cairo in Xenial release and it is not detected by Clair.
% Vuls also missed vulnerabilities such as CVE-2019-5094.

% \textbf{Wrong Detections:} 
% Some vulnerabilities are not linked to Ubuntu 14.04 LTS which is 
% present inside the image, however, Clair have reported such vulnerabilities.
% For instance, CVE-2019-1274 is an example of such vulnerabilities. 
% %is showing because these vulnerabilities are present in Ubuntu 14.04 ESM and 
% %Clair is only referring to Ubuntu 14.04 ESM release. 
% We reported this issue on two Google groups of Clair i.e. CoreOS Dev and
% clair-dev \url{(https://groups.google.com/forum/\#!topic/clair-dev/8IqCCfd-EEc).}

% \textbf{Database difference:} At the time we pulled Clair's database, through \texttt{arminc/clair-db:latest} Docker image, it was one week old.
%                               As a result some vulnerabilities were not updated in the database. For instance, CVE-2019-5094, which
% 		              was published on 2019-09-24, was not 
% 		              updated in Clair's database. Due to which it is missed by Clair.
% 			      Consequently, these discrepancies, even small, are sufficient to change the detection of
% 			      the scanners.

% \textbf{Epoch bug:} There is a bug in the Anchore scanner due to which some vulnerabilities are 
% 		not detected by Anchore. This bug gets triggered when a package’s version have epoch 
% 		(1:3.3.9-1ubuntu2.3) in it. For example, CVE-2018-1125 in Procps package is not
% 		detected by Anchore due to the presence of epoch in the Procps package version.
% 		We reported this bug to the Anchore development team through their Slack channel.

% \textbf{Ubuntu vulnerabilities with status "Ignored (out of standard support)" bug:} There is another bug in Anchore due to 
% 		which it is not able to detect some vulnerabilities. Due to this bug, Anchore does not detect 
% 		vulnerabilities that have a status as “ignored (out of standard support)” in Ubuntu 14.04 LTS. 
% 		However, if this vulnerability is present in Ubuntu 14.04 ESM release, it is detected by Clair. 
% 		For instance, CVE-2019-13565, which is present in Ubuntu 14.04 LTS, is an example of such vulnerabilities. 
% 		We reported this bug as well to the Anchore development team through their Slack
% 		channel.

% \textbf{Rejected CVEs:} There are some CVEs, such as CVE-2017-13753, that are rejected because they are duplicate copies of other CVEs. 
% 	So, Anchore and Clair does not report those rejected CVEs, but Vuls reports them.

% \tristan{integrate this part:

% \change{

% The major discrepancies are shown by Anchore and Clair in Ubuntu \tristan{why?}. Both refers to Ubuntu CVE Tracker database for Ubuntu images.
% The way in which both scanners interpret this database is different.
% %
% %There are different vulnerability databases, which have
% %different formats of data.
% %Out of which Open Vulnerability and Assessment Language (OVAL) is an international
% %open standard and a common
% %means of promoting publicly available security issues. 
% %It provides details about software vulnerabilities. It helps to determine
% %whether vulnerabilities exist in local machines or not. Apart from OVAL, operating system
% %vendors also keep track of their vulnerabilities separately and assign
% %a status to each vulnerability.
% %For example, Ubuntu maintains
% %Ubuntu CVE Tracker, Debian maintains Debian Security Bug Tracker, etc.
% %Each vulnerability is assigned a unique Common Vulnerabilities and Exposures (CVE) number
% %by the MITRE corporation, who maintains a list of all vulnerabilities of all distributions.
% %In this paper, we use vulnerability and CVE interchangeably.
% %Ubuntu CVE Tracker is used by Anchore and Clair in our experiments
% %and there is discrepancies in the way both scanners interpret results.
% %
% %BIDS apps and Boutiques images in our experiments have four Linux
% %distributions i.e. Ubuntu, CentOS, Debian, and Alpine. 
% %Out of these distributions, Ubuntu assigns a vulnerability status to
% %all vulnerabilities before keeping them in its vulnerability database.
% %Each vulnerability is assigned a unique Common Vulnerabilities and Exposures (CVE) number
% %by the MITRE corporation, who maintains a list of all vulnerabilities of all distributions.
% %Ubuntu keeps a record of Ubuntu-specific vulnerabilities in Ubuntu CVE Tracker using their CVE
% %numbers. In this paper, we use vulnerability and CVE interchangeably.
% %The vulnerability status in Ubuntu CVE Tracker is subjected to change
% %depending on the work progress to fix that vulnerability.
% %
% %Regarding Ubuntu, it has different
% %types of releases e.g. Regular, Long Term Support (LTS), and Extended Security Maintenance (ESM) release.
% %Ubuntu maintains its vulnerability data in Ubuntu CVE tracker, which is referred by Anchore and Clair.
% %There is another vulnerability database i.e. Open Vulnerability and Assessment Language (OVAL) repository,
% %which is maintained by MITRE corporation. Vuls refers to OVAL repository for getting Ubuntu vulnerability feed.
% %The vulnerability data format of both databases is different. OVAL keeps a separate xml file for each release,
% %which contains that release specific vulnerability information.
% Ubuntu CVE Tracker keeps a separate file for each vulnerability, which
% has an entry for all Ubuntu releases combined with the current status
% of that vulnerability in that release.
% In particular, Clair and Anchore show discrepancies for Ubuntu 14.04 which have
% two entries for each vulnerability, one for Ubuntu 14.04 Long Term Support (LTS)
% and second for Ubuntu 14.04 Extended Security Maintenance (ESM) release. As Ubuntu 14.04 LTS
% reached end-of-life on 2019-04-30, so Ubuntu ignores new vulnerabilities in LTS release and
% fix them only in ESM release. Consequently, a same vulnerability may have different vulnerability status in LTS and ESM releases,
% which is causing discrepancies. This is because Anchore refers to Ubuntu 14.04 LTS, whereas Clair refers to
% ESM.
% %Further, in Ubuntu a particular release can be of different types i.e. Regular, Long Term Support (LTS), and Extended Security Maintenance (ESM) release. 
% %Through our experiments, we find that Anchore and Clair report Ubuntu vulnerabilities 
% %differently depending on either the different Ubuntu release type or vulnerability status. 
% It is important to provide an overview of these Ubuntu vulnerability statuses, 
% since Ubuntu is heavily utilized in our experiments. 
% %Ubuntu's Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
% %reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
% %in the form of ESM release.
% %ESM is a paid service that can be requested from Ubuntu to get security 
% %updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu version have different vulnerability
% %databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
% %exists in LTS.
% %
% %As an example, Ubuntu has ended support for some releases, such as 14.04 LTS,
% %17.04, and 17.10, as they reached EOL. However, Ubuntu decided to provide optional extended support for Ubuntu 14.04.
% %To use this extended support, users have to purchase Ubuntu 14.04 ESM release.
% %The images in our experiments have Ubuntu 14.04 LTS, so scanners should refer
% %to LTS release only while collecting data from the vulnerability databases.
% %More interestingly, Clair refers to ESM release vulnerability database, however, this may lead to discrepancies in results.
% %Further, if an image is using a Linux release where EOL has been reached that image may
% %turns to be vulnerable as it is not possible to retrieve any updates.
% %
% In Ubuntu CVE Tracker, for a vulnerability all Ubuntu releases are listed with the package name
% in which that vulnerability exists and are entitled a vulnerability status
% (\url{https://git.launchpad.net/ubuntu-cve-tracker/plain/README}), which is
% encoded in the following form:
% \newline \\
% \noindent <release>\_<source-package>: <status> (<version/notes>)
% \newline\\
% The <version/notes> given with a vulnerability status provide more
% information about the vulnerability.
% For a given release, the vulnerability status can be any one of the following:

% \textbf{DNE (Does Not Exist):} The package does not exist in the
% 		release.

% \textbf{needs-triage:} The vulnerability of this package
% 		is not known, and hence evaluation is needed.

% \textbf{not-affected:} The package, while related to the
% 		vulnerability in some way, is not affected by the issue.
% 	%	The <notes> should
% 	%	provide detailed information, if needed.
% 		%For instance, if the given
% 		%status is "not-affected (1.14.6-1.1)" then this indicates that this specific
% 		%package version is not affected.

% \textbf{needed:} The package is vulnerable and needs to be fixed.

% \textbf{active:} The package is vulnerable, needs fixing, and is 
%                  actively being worked on.

% \textbf{ignored:} The package, while related to the
% 		vulnerability in some way, is being ignored for some reason. 
% 	%	The
% 	%	<notes> should provide that reason.
% 	%	For instance, if status looks like
% 	%	"ignored (reached end-of-life)", this means that the given Ubuntu release already reached
% 	%	end-of-life, so this CVE is ignored by Ubuntu. However, if there
% 	%	is extended security support for this release then this
% 	%	CVE will be handled in ESM release.

% \textbf{pending:} The package is vulnerable and
%                   has been fixed but an update has not been yet uploaded or
% 		  published.
% 		 % The <version> is given to indicate the particular 
% 		 % version where the fix has been done.

% \textbf{deferred:} The package is vulnerable, but 
%                    its fix has been deferred for some reason. 
% 		  % The <notes>
% 		  % should provide further details.
% 		  % If a date is mentioned, e.g.
% 		  % "deferred (2015-02-02)", the given date specifies the date when
% 		  % the CVE was deferred.

% \textbf{released:} The package was vulnerable, but
% 		an update has been already uploaded and published.
% 	%	e.g. "released (1.2.3)",
% 	%	<version> indicates the first version where the fix was applied.

% \textbf{released-esm:} The package was vulnerable and
% 		an update has been already uploaded and published. However,
% 		this update is published in the Ubuntu ESM release only and not in LTS.}
% 		%The fixed version of such packages is appended by either
% 		%+esm or \char`\~esm, indicating that this package version is available
% 		%only via ESM.

% 		% \newlne\\
% % Figure~\ref{example} illustrates how the information of a vulnerability is represented
% % in the Ubuntu CVE Tracker.

% % \begin{figure}[!ht]
% % 	\fbox{{\includegraphics[scale=2.5,width=\columnwidth]
% % 	{Figures/vulnExample2.png}}}
% %         \caption{\label{example} Representation of a vulnerability in Ubuntu CVE Tracker}
% % \end{figure}
% % \vspace*{-7mm}
% }

\section{Discussion}

% files vs packages

% reproducibility vs update
However, this action affects the reproducibility
of that image. For example, if a package is used in an experiment
and later that package is updated, subsequent instances of the experiment may actually
get different results.

% minification is specific to the analysis
% tried neurodocker
% don't point at specific applications, it's a general issue

As we noticed that none of the Boutiques and BIDS images at the time of the study are free of vulnerabilities, out of which
so many are high severity vulnerabilities as well. Yet there is not a single image scanner on which we can rely upon for
scanning. Additionally, we cannot update all images because some OS releases, which are used inside images, 
already reached end-of-life and are
no longer getting updates. Furthermore, the minification of images is a time consuming process. As we have to run
the pipeline to get Reprozip trace that can take up to several hours. Also, the user should have proper knowledge
of running the pipeline because it is image-specific. In addition to that, Reprozip can be only installed
in the images having Python version 2.7.3 or greater, or 3.3 or greater. Above all, there is no guarantee that
the commands that you run to get Reprozip trace covers the whole program, maybe it runs a particular
branch of the program. So, there should be some kind of coverage measurement.
Besides this, as scanners report vulnerabilities according to the versions of the installed packages,
they are likely to miss packages that are not installed using package manager. 
Therefore, we suggest some particular rules that can be
followed to keep container images safe.
\begin{itemize}
	\item Update packages regularly inside images.
	\item Remove unnecessary packages, in particular the ones used for installation (compilers, source packages, etc.).
	\item Choose your base OS image wisely.
	\item Run image scanning tools in CI scripts. For instance Anchore.
	\item Protect your GitHub/DockerHub credentials because if automated builds is set up
in Docker Hub, it can automatically build images from external repositories, for instance GitHub.
\item Minimize dependencies inside images.
\item Avoid static linking because it obfuscates vulnerability scanning.
\item Use the package manager to facilitate vulnerability detection.
\item Use OS releases that have LTS support.
\end{itemize}
\vspace*{-2mm}
\section{Conclusions and Future Work}

In this paper, we performed various experiments to study the security state of the Docker
and Singularity images, which are used in the neuroimaging field. We found out strong correlation
between the number of packages and the number of vulnerabilities in the images, which implies
that the number of vulnerabilities increases with the number of packages. We also found out that
updating images reduces the number of vulnerabilities by a significant number. However, for some
vulnerabilities there are no fixes available yet so the images cannot 
be made totally invulnerable. Continuing with our next two experiments, we analyzed that the
image size reduction process also reduces vulnerabilities a lot. However, we found out that this is not
always true. It depends on two factors: the kind of packages and the number of packages that we are
deleting.

In future work we would like to see how these vulnerabilities can be leveraged
to design attacks from containers. Particularly, we want
to consider web servers and clusters that processes data using containers.

\bibliography{bibliography}
\end{document}